On the uniqueness and stability of dictionaries for sparse representation of noisy signals

### 2017-08-15 manuscripts/PNAS/chazthm_pnas.tex
minor updates

% \authordeclaration{The authors declare no conflict of interest.}
\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: chaz@berkeley.edu, chillar@msri.org}

\keywords{Sparse coding, dictionary learning, matrix factorization, compressed sensing, inverse problems, blind source separation}

@@

Note that any particular solution $(\mathbf{B}, \mathbf{\bar x}_1 \ldots, \mathbf{\bar x}_N)$ to this problem gives rise to an orbit of equivalent solutions $(\mathbf{BPD}, \mathbf{D}^{-1}\mathbf{P}^{\top}\mathbf{\bar x}_1, \ldots, \mathbf{D}^{-1}\mathbf{P}^{\top}\mathbf{\bar x}_N)$, where $\mathbf{P}$ is any permutation matrix and $\mathbf{D}$ any invertible diagonal. Previous theoretical work addressing the noiseless case $\eta =0$ (e.g., \cite{li2004analysis, Georgiev05, Aharon06, Hillar15}) has shown that a solution to Prob.~\ref{InverseProblem} (when it exists) is indeed unique up to this inherent ambiguity provided the $\mathbf{x}_i$ are sufficiently diverse and the generating matrix $\mathbf{A}$ satisfies the \textit{spark condition} from compressed sensing:
\begin{align}\label{SparkCondition}
\mathbf{A}\mathbf{x}_1 = \mathbf{A}\mathbf{x}_2 \implies \mathbf{x}_1 = \mathbf{x}_2, \ \ \ \text{for all $k$-sparse } \mathbf{x}_1, \mathbf{x}_2,
\end{align}
%
which would be, in any case, necessary for uniqueness of the $\mathbf{x}_i$. Our concern here is solution stability with respect to noise.

@@

\begin{problem}\label{OptimizationProblem}
Find $\mathbf{B}$ and \mbox{$\mathbf{\bar x}_1, \ldots, \mathbf{\bar x}_N \in \mathbb{R}^m$} that solve:
\begin{align}\label{minsum}
\min \sum_{i = 1}^N \|\mathbf{\bar x}_i\|_0 \ \ 
\text{subject to} \ \ \|\mathbf{z}_i - \mathbf{B}\mathbf{\bar x}_i\|_2 \leq \eta, \ \text{for all $i$}.
\end{align}
\end{problem}
Finally, in Sec.~\ref{Discussion}, we explain how our results fit into the literature and present several applications of our findings.

>>>>

\authordeclaration{The authors declare no conflict of interest.}
\correspondingauthor{\textsuperscript{1}To whom correspondence should be addressed. E-mail: chaz@berkeley.edu, chillar@msri.org}

\keywords{Sparse linear coding, dictionary learning, matrix factorization, compressed sensing, inverse problems, blind source separation}

@@

Note that any particular solution $(\mathbf{B}, \mathbf{\bar x}_1 \ldots, \mathbf{\bar x}_N)$ to this problem gives rise to an orbit of equivalent solutions $(\mathbf{BPD}, \mathbf{D}^{-1}\mathbf{P}^{\top}\mathbf{\bar x}_1, \ldots, \mathbf{D}^{-1}\mathbf{P}^{\top}\mathbf{\bar x}_N)$, where $\mathbf{P}$ is any permutation matrix and $\mathbf{D}$ any invertible diagonal. Previous theoretical work addressing the noiseless case $\eta =0$ (e.g., \cite{li2004analysis, Georgiev05, Aharon06, Hillar15}) has shown that a solution to Prob.~\ref{InverseProblem} (when it exists) is indeed unique up to this inherent ambiguity provided the $\mathbf{x}_i$ are sufficiently diverse and the generating matrix $\mathbf{A}$ satisfies the \textit{spark condition} from compressed sensing:
\begin{align}\label{SparkCondition}
\mathbf{A}\mathbf{x}_1 = \mathbf{A}\mathbf{x}_2 \implies \mathbf{x}_1 = \mathbf{x}_2, \ \ \ \text{for all $k$-sparse } \mathbf{x}_1, \mathbf{x}_2,
\end{align}
%
which would be, in any case, necessary for uniqueness of the $\mathbf{x}_i$. Our concern here is solution stability in the practical setting of noise.

@@

\begin{problem}\label{OptimizationProblem}
Find $\mathbf{B}$ and \mbox{$\mathbf{\bar x}_1, \ldots, \mathbf{\bar x}_N \in \mathbb{R}^m$} that solve:
\begin{align}\label{minsum}
\min \sum_{i = 1}^N \|\mathbf{\bar x}_i\|_0 \ \ 
\text{subject to} \ \ \|\mathbf{z}_i - \mathbf{B}\mathbf{\bar x}_i\|_2 \leq \eta, \ \text{for all $i$}.
\end{align}
\end{problem}
Finally, in Sec.~\ref{Discussion}, we present several applications of our findings.
